---
title: 'Code for: Impact of preexisting dengue immunity on Zika virus emergence in a dengue endemic region'
author: "Rodriguez-Barraquer I, Costa F. et al."
output:
  html_document: default
  html_notebook: default
---

This is file contains the code to reproduce the main results in the paper "Impact of preexisting dengue immunity on Zika virus emergence in a dengue endemic region".

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
expit.fun<-function(x) {exp(x)/(1+exp(x))}
```

```{r load_data}
#dataset
db<-readRDS("../Data/dat_Zika_salvador.RDS")
cut1<-0.585592 #Cut-point determined using validation set
db$pos_infect_38<-db$pos_igg3<-ifelse(db$avgzikvigg3_l38>cut1, 1, 0)
db$coh_642<-ifelse(!is.na(db$denviggt_titers_l37_new), 1, 0)
```

## Estimate proportion seropositive prior to the ZIKV outbreak
Samples belong to three different collection periods
-L36 October 2014
-L37 March 2015
-L38 October 2015

### By IgG3
```{r seropos_pre_igg3_l36}
db$pos_infect_36<-ifelse(db$avgzikvigg3_l36>=cut1, 1, 0)
#prop.table(table(db$pos_infect_36))

n.boot<-1000 #Number of bootstrap iterations for confidence intervals
n.pos<-c()

for(i in 1:n.boot){
  samp.i<-sample(db$pos_infect_36, size = 1453, replace = TRUE)
  n.pos[i]<-sum(samp.i, na.rm=T)
}
#quantile(n.pos, c(.025, .975))/length(which(!is.na(db$pos_infect_36)))
#length(which(!is.na(db$pos_infect_36)))
```
Number of anti-ZIKV NS1 IgG3 positive samples in October 2014: `r length(which(db$pos_infect_36==1))`/`r length(which(!is.na(db$pos_infect_36)))`, `r paste(round(prop.table(table(db$pos_infect_36))[2],2), " (95%CI ", paste(round(quantile(n.pos, c(.025, .975))/length(which(!is.na(db$pos_infect_36))), 2), collapse="-"), ")", sep="" )`


```{r seropos_pre_igg3_l37}
db$pos_infect_37<-ifelse(db$avgzikvigg3_l37>=cut1, 1, 0)
#prop.table(table(db$pos_infect_37))
for(i in 1:n.boot){
  samp.i<-sample(db$pos_infect_37, size = 1453, replace = TRUE)
  n.pos[i]<-sum(samp.i, na.rm=T)
}
#quantile(n.pos, c(.025, .975))/length(which(!is.na(db$pos_infect_37)))
#length(which(!is.na(db$pos_infect_37)))
```
Number of anti-ZIKV NS1 IgG3 positive samples in March 2014: `r length(which(db$pos_infect_37==1))`/`r length(which(!is.na(db$pos_infect_37)))`, `r paste(round(prop.table(table(db$pos_infect_37))[2],2), " (95%CI ", paste(round(quantile(n.pos, c(.025, .975))/length(which(!is.na(db$pos_infect_37))), 2), collapse="-"), ")", sep="" )`


### By PRNT
```{r seropos_pre_prnt}
### Same but by PRNT
db$prnt_37<-ifelse(db$log10zikv_prnt37_50> 2, 1, 0)
#prop.table(table(db$prnt_37))
for(i in 1:n.boot){
  samp.i<-sample(db$prnt_37, size = 1453, replace = TRUE)
  n.pos[i]<-sum(samp.i, na.rm=T)
}
#quantile(n.pos, c(.025, .975))/length(which(!is.na(db$pos_infect_37)))
#length(which(!is.na(db$prnt_37)))
```
Number of ZIKV PRNT positive samples in March 2014: `r length(which(db$prnt_37==1))`/`r length(which(!is.na(db$prnt_37)))`, `r paste(round(prop.table(table(db$prnt_37))[2],2), " (95%CI ", paste(round(quantile(n.pos, c(.025, .975))/length(which(!is.na(db$prnt_37))), 2), collapse="-"), ")", sep="" )`


## Estimate attack rates of ZIKV outbreak

```{r attack_rates_igg3}


### Formula to adjust seropositivity based on known sensitivity and specificity
form.correct<-function(sp, se, pos, neg) {((sp*pos-(1-sp)*neg)/(se*sp-(1-se)*(1-sp)))}

adj.rate<-form.correct(sp =0.9722222, se = 0.85245902, pos =table(db$pos_infect_38)[2], neg = table(db$pos_infect_38)[1]) 

# Bootstrap to obtain confidence bounds

n.boot<-1000
n.pos<-c()

for(i in 1:n.boot){
  samp.i<-sample(db$pos_infect_38[which(!is.na(db$pos_infect_38))], size = 1453, replace = TRUE)
  n.pos[i]<-sum(samp.i)
}

adj.rate<-form.correct(sp =0.9722222, se = 0.85245902, pos =n.pos, neg = 1453-n.pos )/1453
#quantile(adj.rate, c(.025, .975))
```
Number of anti-ZIKV NS1 IgG3 positive samples in October 2015: `r length(which(db$pos_infect_38==1))`/`r length(which(!is.na(db$pos_infect_38)))`, `r paste(round(prop.table(table(db$pos_infect_38))[2],2), " (95%CI ", paste(round(quantile(n.pos, c(.025, .975))/length(which(!is.na(db$pos_infect_38))), 2), collapse="-"), ")", sep="" )`

Attack rate adjusted for sensitivity and specificity: `r paste(round(mean(adj.rate), 2), " (95%CI ", paste(round(quantile(adj.rate, c(.025, .975)), 2), collapse="-"), ")", sep="") `

### Look at spatial heterogeneity in attack rates
These variables not provided due to confidentiality concerns

## Look at associations between prior immunity to DENV and ZIKV seropositivity
These analyses are performed on subset of 642 with data on DENV titers prior to Zika outbreak (variable coh_642=1)

### Association among individuals with prior immunity to DENV
```{r associations_DENV_infect}
mod.all_pos<-glm(pos_igg3~log(avgdenvigg3_l37, 2)+idade_l38+log(denviggt_titers_l37_new, 2), data=db[which(db$denviggt_titers_l37_new>50),], family="binomial")
summary(mod.all_pos)

1-exp(coef(mod.all_pos))
1-exp(coef(mod.all_pos)+ 1.96*summary(mod.all_pos)$coefficients[,2] )
1-exp(coef(mod.all_pos)- 1.96*summary(mod.all_pos)$coefficients[,2] )

```


### Association between ZIKV infection and tertiles of DENV titers
```{r associations_DENV_infect_tertiles}
db$cut_denv_iggt_new<-cut(db$denviggt_titers_l37_new, quantile(db$denviggt_titers_l37_new, c(0, 1/3, 2/3, 3/3), na.rm=T), include.lowest = T)
### Alternative, comparing to now titers -
#db$cut_denv_iggt_new<-cut(db$denviggt_titers_l37_new, c(0, 51, quant.denv), include.lowest = T)
table(db$cut_denv_iggt_new, useNA = 'always')

mod.all_quants<-glm(pos_igg3~log(avgdenvigg3_l37, 2)+idade_l38+cut_denv_iggt_new, data=db, family="binomial")
summary(mod.all_quants)
1-exp(coef(mod.all_quants))
1-exp(coef(mod.all_quants)+ 1.96*summary(mod.all_quants)$coefficients[,2] )
1-exp(coef(mod.all_quants)- 1.96*summary(mod.all_quants)$coefficients[,2] )

### Same but with quasipoisson regression to get PRR instead of OR
#mod.all_quants<-glm(pos_igg3~log(avgdenvigg3_l37, 2)+idade_l38+cut_denv_iggt_new, data=db, family=quasipoisson(link = "log"))
#summary(mod.all_quants)
#1-exp(coef(mod.all_quants))
#1-exp(coef(mod.all_quants)+ 1.96*summary(mod.all_quants)$coefficients[,2] )
#1-exp(coef(mod.all_quants)- 1.96*summary(mod.all_quants)$coefficients[,2] )
#dim(mod.all_quants$model)
```

## Look at associations between prior immunity to DENV and ZIKV symptoms

### First look at proportion of individuals who presented symptoms and association between symptoms and infection

```{r symptoms, echo=TRUE}
###### Symptoms
table(db$rash_l38[which(db$coh_642==1)])
prop.table(table(db$rash_l38[which(db$coh_642==1)]))
table(db$febre_l38[which(db$coh_642==1)])
prop.table(  table(db$febre_l38[which(db$coh_642==1)]))
table(db$rash_l38[which(db$coh_642==1)], db$febre_l38[which(db$coh_642==1)])
prop.table(table(db$rash_l38[which(db$coh_642==1)], db$febre_l38[which(db$coh_642==1)]))

db$symp<-ifelse(db$rash_l38==1 | db$febre_l38==1, 1, 0)
db$symp2<-ifelse(db$rash_l38==1 & db$febre_l38==1, 1, 0)
table(db$symp, db$pos_infect_38)
prop.table(table(db$symp, db$pos_infect_38), 2)

db$symp<-ifelse(db$rash_l38==1 | db$febre_l38==1, 1, 0)
db$symp2<-ifelse(db$rash_l38==1 & db$febre_l38==1, 1, 0)
table(db$symp, db$pos_infect_38)
prop.table(table(db$symp, db$pos_infect_38), 2)

db$rash2<-ifelse(db$rash_l38==1, 1, 0)

## Association between infection and rash
m.rash<-glm(pos_infect_38~rash2, family="binomial", data=db[which(db$coh_642==1),])
exp(coef(m.rash))
exp(coef(m.rash)+ 1.96*summary(m.rash)$coefficients[,2] )
exp(coef(m.rash)- 1.96*summary(m.rash)$coefficients[,2] )

## Association between infection and fever
m.fev<-glm(pos_infect_38~febre_l38, family="binomial", data=db[which(db$coh_642==1),])
exp(coef(m.fev))
exp(coef(m.fev)+ 1.96*summary(m.fev)$coefficients[,2] )
exp(coef(m.fev)- 1.96*summary(m.fev)$coefficients[,2] )

## Association between infection and both symptoms
m.symp<-glm(pos_infect_38~symp2, family="binomial", data=db[which(db$coh_642==1),])
exp(coef(m.symp))
exp(coef(m.symp)+ 1.96*summary(m.symp)$coefficients[,2] )
exp(coef(m.symp)- 1.96*summary(m.symp)$coefficients[,2] )

```

### Now look at association between DENV titers and symptoms among individuals who were infected

#### Fever
```{r Fever_DENV, echo=FALSE}
m.denv_febre<-glm(febre_l38~cut_denv_iggt_new+log(avgdenvigg3_l37,2)+idade_l38, family="binomial", data=db[which(db$coh_642==1 & db$pos_infect_38==1),])
summary(m.denv_febre)

```

#### Rash
```{r rash_DENV, echo=FALSE}
m.denv_rash<-glm(rash2~cut_denv_iggt_new+log(avgdenvigg3_l37, 2)+idade_l38, family="binomial", data=db[which(db$coh_642==1 & db$pos_infect_38==1),])


summary(m.denv_rash)

```


